Cell Decompositions for Rank Two Quiver Grassmannians

### 2017-12-21 Grassmannian_cells.tex
startetd new proof of Lemma 3.3

\begin{lemma}
  \label{le:injective evaluation maps}
  For any $V\in Gr(H_m)$, the natural evaluation map $ev_V:P_m\otimes V\to P_{m+1}$ is injective.
\end{lemma}
%%%%%%%%%%%%%%%%%%%
% needs polishing %
%%%%%%%%%%%%%%%%%%%
\begin{proof}
  \footnote{This needs to be cleaned up, the proof seems false as written.}
  A nonzero element $f\in V$ corresponds to an irreducible morphism $f:P_m\to P_{m+1}$ and thus does not factor through any representation $Z\ne P_m$, in particular $f$ is injective.
  But any nonzero element of $\ker ev_V$ gives rise to a nonzero $f\in V$ which is not injective and thus $ev_V$ must be an injective map.
\end{proof}
In what follows we will not distinguish between $P_m\otimes V$ and its image under $ev_V$.

@@

For each $m\ge2$, there is an Auslander-Reiten sequence \cite{auslander-reiten-smalo}
\begin{equation}
  \label{eq:AR sequence}
  0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m\otimes H_m\stackrel{ev}{\longrightarrow} P_{m+1}\longrightarrow 0,
\end{equation}
where the right-hand morphism is the natural evaluation map.
\begin{corollary}
  \label{cor:preprojective homomorphism duality}
  The sequence \eqref{eq:AR sequence} gives rise to an isomorphism $H_m^*\cong H_{m-1}$.
\end{corollary}
\begin{proof}
  Define a map $H_m^*\to H_{m-1}$ by $\varphi\mapsto\bar{\varphi}:=(id\overline{\otimes}\varphi)\circ\iota_{m-1}$, in words $\bar{\varphi}$ acts on $x\in P_{m-1}$ by contracting with the second factor in $\iota_{m-1}(x)$ to give an element of $P_m$.
  Suppose $\varphi\in H_m^*$ is a nonzero functional on $H_m$ and let $V\subsetneq H_m$ denote the kernel of $\varphi$.
  Then $\bar{\varphi}=0\in H_{m-1}$ if and only if the image of $\iota_{m-1}$ is contained in $P_m\otimes V\subsetneq P_m\otimes H_m$.
  But then Lemma~\ref{le:injective evaluation maps} implies $\iota_{m-1}$ is the zero map, a contradiction.
  Thus the map $H_m^*\to H_{m-1}$, $\varphi\mapsto\bar{\varphi}$ must be injective and hence an isomorphism.
\end{proof}
\begin{remark}
  The set $Gr(H_m)$ is naturally a poset and Corollary~\ref{cor:preprojective homomorphism duality} gives an identification of the opposite poset $Gr(H_m)^{op}\cong Gr(H_m^*)$ with $Gr(H_{m-1})$.
  We write $\bar{V}\subset H_{m-1}$ for the subspace corresponding to $V\subset H_m$ under this identification.
  Under the isomorphism of $H_{m-1}$ with $H_m^*$, we have $\bar{V}=(H_m/V)^*$.
\end{remark}
Following Remark~\ref{rem:reflection recursion}, the reflection functor $\Sigma_1$ provides a natural identification of $H_m$ with $H_{m-1}$ for $m\ge2$. 
Combining with Corollary~\ref{cor:preprojective homomorphism duality} we obtain a natural isomorphism $H_m^*\cong H_m$ for $m\ge2$.
The next observation is essential for the results to follow.
\begin{proposition}
  For $V\in Gr(H_m)$, $m\ge1$, we have $\bar{\bar{V}}=\tau V$.
\end{proposition}
\begin{proof}
  We show that $\bar{\bar{\varphi}}=\tau\varphi$ for $\varphi\in H_m$.
  Indeed, we have
  \[\bar{\bar{\varphi}}=(id\overline{\otimes}\bar{\varphi})\circ\iota_{m-2}=\Big(id\overline{\otimes}\big((id\overline{\otimes}\varphi)\circ\iota_{m-1}\big)\Big)\circ\iota_{m-2}.\]
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} S. Benner, M.~C.~R. Butler, The equivalence of certain functors occuring in the representation theory of artin algebras and species, J. London Math. Soc., 14 (1976), 183-187.
\bibitem{cc} P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck}
  P. Caldero, B. Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{cefr} Cerulli Irelli, Esposito, Franzen, Reineke: Topology of Quiver Grassmannians. Preprint 2017.
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).

>>>>

For each $m\ge2$, there is an Auslander-Reiten sequence \cite[Section V]{ars}
\begin{equation}
  \label{eq:AR sequence}
  0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m\otimes H_m\stackrel{ev}{\longrightarrow} P_{m+1}\longrightarrow 0,
\end{equation}
where the right-hand morphism is the natural evaluation map.

\begin{lemma}
  \label{le:injective evaluation maps}
  For any $V\in Gr(H_m)$, the natural evaluation map $ev_V:P_m\otimes V\to P_{m+1}$ is injective.
\end{lemma}
%%%%%%%%%%%%%%%%%%%
% needs polishing %
%%%%%%%%%%%%%%%%%%%
\begin{proof}
As $\Ext(P_m\otimes V,P_{m-1})=0$, we obtain a commutative diagram
\[\xymatrix{0 \ar[r] & P_{m-1}\ar[r]^{\iota_{m-1}} & P_m\otimes H_m \ar[r]^{ev} &P_{m+1} \ar[r] & 0\\0 \ar[r] & P_{m-1}\ar@{=}[u]\ar[r] &P_{m-1}\oplus P_m\otimes V \ar[r]\ar[u]^{(\iota_{m-1},\mathrm{id}_{P_m}\otimes \iota_V)} &P_{m}\otimes V \ar[r]\ar^{ev_V}[u] & 0}.\]

% \footnote{This needs to be cleaned up, the proof seems false as written.}
 % A nonzero element $f\in V$ corresponds to an irreducible morphism $f:P_m\to P_{m+1}$ and thus does not factor through any representation $Z\ne P_m$, in particular $f$ is injective.
 % But any nonzero element of $\ker ev_V$ gives rise to a nonzero $f\in V$ which is not injective and thus $ev_V$ must be an injective map.
\end{proof}
In what follows we will not distinguish between $P_m\otimes V$ and its image under $ev_V$.

@@

\begin{corollary}
  \label{cor:preprojective homomorphism duality}
  The sequence \eqref{eq:AR sequence} gives rise to an isomorphism $H_m^*\cong H_{m-1}$.
\end{corollary}
\begin{proof}
  Define a map $H_m^*\to H_{m-1}$ by $\varphi\mapsto\bar{\varphi}:=(id\overline{\otimes}\varphi)\circ\iota_{m-1}$, in words $\bar{\varphi}$ acts on $x\in P_{m-1}$ by contracting with the second factor in $\iota_{m-1}(x)$ to give an element of $P_m$.
  Suppose $\varphi\in H_m^*$ is a nonzero functional on $H_m$ and let $V\subsetneq H_m$ denote the kernel of $\varphi$.
  Then $\bar{\varphi}=0\in H_{m-1}$ if and only if the image of $\iota_{m-1}$ is contained in $P_m\otimes V\subsetneq P_m\otimes H_m$.
  But then Lemma~\ref{le:injective evaluation maps} implies $\iota_{m-1}$ is the zero map, a contradiction.
  Thus the map $H_m^*\to H_{m-1}$, $\varphi\mapsto\bar{\varphi}$ must be injective and hence an isomorphism.
\end{proof}
\begin{remark}
  The set $Gr(H_m)$ is naturally a poset and Corollary~\ref{cor:preprojective homomorphism duality} gives an identification of the opposite poset $Gr(H_m)^{op}\cong Gr(H_m^*)$ with $Gr(H_{m-1})$.
  We write $\bar{V}\subset H_{m-1}$ for the subspace corresponding to $V\subset H_m$ under this identification.
  Under the isomorphism of $H_{m-1}$ with $H_m^*$, we have $\bar{V}=(H_m/V)^*$.
\end{remark}
Following Remark~\ref{rem:reflection recursion}, the reflection functor $\Sigma_1$ provides a natural identification of $H_m$ with $H_{m-1}$ for $m\ge2$. 
Combining with Corollary~\ref{cor:preprojective homomorphism duality} we obtain a natural isomorphism $H_m^*\cong H_m$ for $m\ge2$.
The next observation is essential for the results to follow.
\begin{proposition}
  For $V\in Gr(H_m)$, $m\ge1$, we have $\bar{\bar{V}}=\tau V$.
\end{proposition}
\begin{proof}
  We show that $\bar{\bar{\varphi}}=\tau\varphi$ for $\varphi\in H_m$.
  Indeed, we have
  \[\bar{\bar{\varphi}}=(id\overline{\otimes}\bar{\varphi})\circ\iota_{m-2}=\Big(id\overline{\otimes}\big((id\overline{\otimes}\varphi)\circ\iota_{m-1}\big)\Big)\circ\iota_{m-2}.\]
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{ars} Auslander, M., Reiten, I., Smalo S.O.: Representation theory of Artin algebras {\bf 36}. Cambridge University Press, Cambridge 1997.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} S. Benner, M.~C.~R. Butler, The equivalence of certain functors occuring in the representation theory of artin algebras and species, J. London Math. Soc., 14 (1976), 183-187.
\bibitem{cc} P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck}
  P. Caldero, B. Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{cefr} Cerulli Irelli, Esposito, Franzen, Reineke: Topology of Quiver Grassmannians. Preprint 2017.
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).
